Computable Poincar\'e--Friedrichs constants\\ for the $L^{p}$ de Rham complex over domains\\ with shellable triangulations

### 2024-09-03 main.tex
last sections

\begin{theorem}\label{theorem:starreflection}
    Let $\calT$ be an $n$-dimensional triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be a simplex of dimension $k < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        U := |\patch_{\calT}(S)| \setminus T.
    \end{align*}
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi : T \rightarrow U
    \end{align*}
    satisfying
    \begin{align*}
        \Lip( \Xi,      T ) \leq \Cvier{n}{k}(\calT),
        \quad 
        \Lip( \Xi^{-1}, U ) \leq \Cfive{n}{k}(\calT),
    \end{align*}
    where $\Cvier{n}{k}(\calT)$ and $\Cfive{n}{k}(\calT)$ are constants as stated in the proof. 
\end{theorem}
\begin{proof}
    We derive the estimate in several steps. 
    \begin{itemize}
        \item 
        Without loss of generality, the barycenter of $S$ is the origin. 
        % Let $I := U \cap T$, which is the union of $k$ different faces of $T$, with $1 \leq k \leq n$.
        %We let $J \subseteq \partial T$ be the union of the remaining $n+1-k$ faces of $T$.
        %So $J$ is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$.
        Note that $J$ is the union of exactly those faces of $T$ that contain $S'$.

@@

        \item 
        We introduce a new mapping $\Theta : |\calR| \rightarrow |\calR^c|$.
        Let $K \in \calR$ be an $n$-simplex and $K' \in \calR^{c}$ be constructed from $K$.
        We let $A_{K} : \Delta^{n} \rightarrow K$ and $A_{K'} : \Delta^{n} \rightarrow K'$
        be affine reference transformations
        that agree on the vertices common to $K$ and $K'$.
        We then define 
        \begin{align*}
            \Theta_{|K} := A_{K'} \circ A_{K}^{-1}.
        \end{align*}
        Notice that $A_{K'}$ and $A_{K}$ have the same singular values. Thus
        \begin{align*}
            \| \Jacobian \Theta_{|K}      \| &\leq \| A_{K'} \| \cdot \| A_{K}^{-1} \| \leq \kappa(A_{K}),
            \\
            \| \Jacobian \Theta_{|K}^{-1} \| &\leq \| A_{K'}^{-1} \| \cdot \| A_{K} \| \leq \kappa(A_{K}).
        \end{align*}

@@

\begin{lemma}\label{lemma:mixedbconsimplex:exteriorderivative}
    Let $T$ be an $n$-simplex 
    and let $F_{0},\dots,F_{l}$ be $l+1$ different faces of $T$. 
    Suppose that $u \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} u = 0$ for $0 \leq i \leq l$.
    Then there exists $w \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} w = 0$ for $0 \leq i \leq l$
    and  
    \begin{align*}
        \cartan u = \cartan w,
        \quad 
        \| w \| 
        \leq 
        C_{{\rm PF},T,F,k,p}
        \| \cartan u \|
        ,
    \end{align*}
    where $C_{{\rm PF},T,F,k,p} > 0$ is a constant such that 
    \begin{align*}
        C_{{\rm PF},T,F,k,p}
        \leq 
        ...
        \diam(T)
        .
    \end{align*}
\end{lemma}
\begin{proof}
    There exists $\varphi : \hat T \rightarrow T_{m+1}$ mapping the convex closure of the $n$ unit vectors onto the face $F_0$.
    We can also assume that the face of $\hat T$
    orthogonal to the $i$-th coordinate axis is mapped 
    onto the face $F_i$. 
    In what follows, we let $\tilde U$ be the domain obtained from reflecting $\hat T$ along the coordinate axes $l+1$ through $n$.

@@

    The bounds for the potential operators imply 
    the existence of $\tilde w \in W^{p}_{0}\Lambda^{k}(\tilde U)$
    with $\cartan \tilde w = \cartan \tilde u$ and 
    \begin{align*}
        \| \tilde w \|_{L^{p}(\tilde U)}
        &\leq 
        n A_{\Bogov}(n,k+1) \frac{ \vol_{n}(B_2(0)) }{|\Omega|} 
        2
        \| \tilde g \|_{L^{p}(\tilde U)}
        \\&\leq 
        2^{n+1} n A_{\Bogov}(n,k+1) \frac{ \vol_{n}(B_1(0)) }{|\Omega|} 
        \| \tilde g \|_{L^{p}(\tilde U)}
        .
    \end{align*}
    We let $w \in W^{p}\Lambda^{k}(T)$ be defined by $w := \varphi^{-\ast} \tilde w$.
    By construction, $\cartan w = \cartan u$,
    and $w$ has vanishing trace along the faces $F_{0},\dots,F_{l}$.

    In addition to that, by Theorem~\ref{theorem:pullbackestimate}:
    \begin{gather*}
        \| w \|_{L^{p}(T_{m+1})}
        \leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \| \tilde w \|_{L^{p}(\hat T)}
        \leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \| \tilde w \|_{L^{p}(\tilde U)}
        ,
        \\
        \| \tilde g \|_{L^{p}(\tilde U)}
        \leq 
        2^{\frac n p}
        \| \hat g \|_{L^{p}(\hat T)}
        \leq 
        2^{\frac n p}
        |\det( \Jacobian \varphi )|^{-\frac 1 p} \| \Jacobian \varphi \| \cdot 
        \| \cartan u \|_{L^{p}(T)}
        .
    \end{gather*}
    The second inequality follows.
\end{proof}

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    and such that for all $T \in \calT$:
    \begin{align*}
        \| w \|_{L^{p}(T_m)}
        \leq 
        ...
    \end{align*}
\end{theorem}

@@

    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    %Suppose that claim is true when the domain has $m$ simplices of dimension $n$. 
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(\Omega_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $\Omega_{m-1}$. 
    By assumption, $T_{m}$ and $\overline\Omega_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex, and we let $U_{m} \subseteq \Omega_{m}$ be the complement of $T_{m}$ in that newly completed star.

@@

    We want to estimate norms. 
    By construction,
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_{m})}
        \leq  
        \| w'_{m} \|_{L^{p}(T_{m})}
        + 
        \| w''_{m} \|_{L^{p}(T_{m})}
        .
    \end{align*}
    Due to Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative},
    we can assume 
    \begin{align*}
        \| w''_{m} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} \| \cartan u - \cartan w'_{m} \|_{L^{p}(T_{m})}
        \\&
        \leq 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} \| \cartan u      \|_{L^{p}(T_{m})} 
        + 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} \| \cartan w'_{m} \|_{L^{p}(T_{m})} 
        .
    \end{align*}
    We also employ Theorem~\ref{theorem:pullbackestimate} to find 
    \begin{align*}
        \| w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \| \Jacobian \Xi \|^{k}
        \cdot 
        \| w_{m-1} \|_{L^{p}(U_{m-1})}
        ,
        \\
        \| \cartan w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \| \Jacobian \Xi \|^{k+1}
        \cdot 
        \| \cartan w_{m-1} \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    We have assumed that $\cartan w_{m-1} = \cartan u_{|\Omega_{m-1}}$. 
    By the properties of the reflection mapping in Theorem~\ref{theorem:starreflection},
    \begin{align*}
        ...
    \end{align*}
    This enables the recursion
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_{m})}
        &
        \leq  
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \| \Jacobian \Xi \|^{k} 
        \| w_{m-1} \|_{L^{p}(U_{m-1})}
        \\&\qquad
        + 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} 
        \| \cartan u      \|_{L^{p}(T_{m})} 
        \\&\qquad
        + 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} 
        |\det(\Jacobian \Xi  )|^{-\frac 1 p} 
        \| \Jacobian \Xi \|^{k+1} 
        \| \cartan u_{m-1} \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    The existence of $w \in W^{p}\Alt^{k}(\Omega)$ satisfying the recursive estimate follows. 
\end{proof}

>>>>

\begin{theorem}\label{theorem:starreflection}
    Let $\calT$ be an $n$-dimensional triangulation of an $n$-dimensional domain. 
    Let $S \in \calT$ be a simplex of dimension $k < n$,
    let $T \in \patch_{\calT}(S)$ be of dimension $n$,
    and let 
    \begin{align*}
        U := |\patch_{\calT}(S)| \setminus T.
    \end{align*}
    There exists a bi-Lipschitz mapping
    \begin{align*}
        \Xi : T \rightarrow U
    \end{align*}
    satisfying
    \begin{gather*}
        \| \Jacobian\Xi \|          \leq \Cvier{n}(\calT),
        \quad 
        \| \Jacobian\Xi^{-1} \|     \leq \Cfive{n}(\calT),
        \\
        \| \det\Jacobian\Xi \|      \leq \Cdetvier{n}(\calT),
        \quad 
        \| \det\Jacobian\Xi^{-1} \| \leq \Cdetfive{n}(\calT),
    \end{gather*}
    where the constants on the right-hand sides are as stated in the proof. 
\end{theorem}
\begin{proof}
    We derive the estimate in several steps. 
    \begin{itemize}
        \item 
        Without loss of generality, the barycenter of $S$ is the origin. 
        % Let $I := U \cap T$, which is the union of $k$ different faces of $T$, with $1 \leq k \leq n$.
        %We let $J \subseteq \partial T$ be the union of the remaining $n+1-k$ faces of $T$.
        %So $J$ is within the boundary of the star $|\patch_{\calT}(S)|$ and is disjoint from $S$.
        We fix the subsimplex $S' \subseteq T$ that is complementary to $S$.
        Note that $J$ is the union of exactly those faces of $T$ that contain $S'$.

@@

        \item 
        We introduce a new mapping $\Theta : |\calR| \rightarrow |\calR^c|$.
        Let $K \in \calR$ be an $n$-simplex and $K' \in \calR^{c}$ be constructed from $K$.
        We let $A_{K} : \Delta^{n} \rightarrow K$ and $A_{K'} : \Delta^{n} \rightarrow K'$
        be affine reference transformations
        that agree on the vertices common to $K$ and $K'$.
        We then define 
        \begin{align*}
            \Theta_{|K} := A_{K'} \circ A_{K}^{-1}.
        \end{align*}
        Notice that $A_{K'}$ and $A_{K}$ have the same singular values. Thus
        \begin{align*}
            \| \Jacobian \Theta_{|K}      \| &\leq \| A_{K'} \| \cdot \| A_{K}^{-1} \| \leq \kappa(A_{K}),
            \\
            \| \Jacobian \Theta_{|K}^{-1} \| &\leq \| A_{K'}^{-1} \| \cdot \| A_{K} \| \leq \kappa(A_{K}).
        \end{align*}
        \color{red}
        By construction, 
        \begin{align*}
            \det( \Jacobian \Theta_{|K} )
            = 
            \det( \Jacobian^{-1} \Theta_{|K} )
            =
            1
            .
        \end{align*}
        \color{black}

@@

\begin{lemma}\label{lemma:mixedbconsimplex:exteriorderivative}
    Let $T$ be an $n$-simplex 
    and let $F_{0},\dots,F_{l}$ be $l+1$ different faces of $T$. 
    Suppose that $u \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} u = 0$ for $0 \leq i \leq l$.
    Then there exists $w \in W^{p}\Lambda^{k}(T)$ such that 
    $\trace_{T,F_i} w = 0$ for $0 \leq i \leq l$
    and  
    \begin{align*}
        \cartan u = \cartan w,
        \quad 
        \| w \| 
        \leq 
        C_{{\rm PF},\Gamma,l,k,p}(T)
        \diam(T)^{k} 
        \| \cartan u \|
        .
    \end{align*}
    Here, $C_{{\rm PF},\Gamma,l,k,p}(T) > 0$ is a constant such that 
    \begin{align*}
        C_{{\rm PF},\Gamma,l,k,p}(T)
        \leq 
        2^{\frac n p}
        2^{n+1} n A_{\Bogov}(n,k+1) \frac{ \vol_{n}(B_1(0)) }{ (l+1)/n! } 
        \Ceins{n}(T)^{k}
        % \| \Jacobian \varphi \|^{k}
        .
    \end{align*}
\end{lemma}
\begin{proof}
    There exists $\varphi : \hat T \rightarrow T_{m+1}$ mapping the convex closure of the $n$ unit vectors onto the face $F_0$.
    We can also assume that the face of $\hat T$
    orthogonal to the $i$-th coordinate axis is mapped 
    onto the face $F_i$. 
    In what follows, we let $\tilde U$ be the domain obtained from reflecting $\hat T$ along the coordinate axes $l+1$ through $n$. 
    We see that $\vol(\tilde U) = (l+1)/n!$

@@

    The bounds for the potential operators imply 
    the existence of $\tilde w \in W^{p}_{0}\Lambda^{k}(\tilde U)$
    with $\cartan \tilde w = \cartan \tilde u$ and 
    \begin{align*}
        \| \tilde w \|_{L^{p}(\tilde U)}
        &\leq 
        n A_{\Bogov}(n,k+1) \frac{ \vol_{n}(B_2(0)) }{ \vol(\tilde U) } 
        2
        \| \tilde g \|_{L^{p}(\tilde U)}
        \\&\leq 
        2^{n+1} n A_{\Bogov}(n,k+1) \frac{ \vol_{n}(B_1(0)) }{ (l+1)/n! } 
        \| \tilde g \|_{L^{p}(\tilde U)}
        .
    \end{align*}
    We let $w \in W^{p}\Lambda^{k}(T)$ be defined by $w := \varphi^{-\ast} \tilde w$.
    By construction, $\cartan w = \cartan u$,
    and $w$ has vanishing trace along the faces $F_{0},\dots,F_{l}$.

    In addition to that, by Theorem~\ref{theorem:pullbackestimate}:
    \begin{gather*}
        \| w \|_{L^{p}(T_{m+1})}
        \leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \| \tilde w \|_{L^{p}(\hat T)}
        \leq 
        |\det(\Jacobian \varphi)|^{\frac 1 p} 
        \| \tilde w \|_{L^{p}(\tilde U)}
        ,
        \\
        \| \tilde g \|_{L^{p}(\tilde U)}
        \leq 
        2^{\frac n p}
        \| \hat g \|_{L^{p}(\hat T)}
        \leq 
        2^{\frac n p}
        |\det( \Jacobian \varphi )|^{-\frac 1 p} 
        \| \Jacobian \varphi \|^{k}
        \cdot 
        \| \cartan u \|_{L^{p}(T)}
        .
    \end{gather*}
    The second inequality follows.
\end{proof}

In what follows, use the abbreviation
\begin{align*}
    C_{{\rm PF},\Gamma,k,p}(\calT) 
    :=
    \max\limits_{ \substack{ T \in \calT \; \dim(T)=n \\ 0 \leq l \leq n } }
    C_{{\rm PF},\Gamma,l,k,p}(T)   
\end{align*}
whenever $\calT$ is an $n$-dimensional triangulation.

@@

\begin{theorem}\label{theorem:poincarefriedrichsestimate:exterior}
    Let $\Omega \subseteq \bbR^{n}$ be a domain with a shellable triangulation $\calT$.
    Let $T_0, T_1, \dots, T_m$ be a shelling.
    Then for any $u \in W^{p}\Alt^{k}(\Omega)$ 
    there exists $w \in W^{p}\Alt^{k}(\Omega)$ with $\cartan u = \cartan w$ 
    and such that for each $T_{m} \in \calT$ in the shelling
    we have the following recursive estimate: 
    \begin{align*}
        \| w \|_{L^{p}(T_{m})}
%         &
%         \leq  
%         C_{{\rm PF},\Gamma,k,p}(\calT) 
%         \Cdetfive{n}(\calT)^{\frac 1 p} 
%         \Cvier{n}(\calT)^{k+1} 
%         \diam(T)^{k} 
%         \| \cartan u \|_{L^{p}(U_{m-1})}
%         \\&\qquad
%         + 
%         C_{{\rm PF},\Gamma,k,p}(\calT) 
%         \diam(T)^{k} 
%         \| \cartan u      \|_{L^{p}(T_{m})} 
%         \\&\qquad
%         + 
%         \Cdetfive{n}(\calT)^{-\frac 1 p} 
%         \Cvier{n}(\calT)^{k} 
%         \| w \|_{L^{p}(U_{m-1})}
%         \\
        &
        \leq  
        C_{{\rm PF},\Gamma,k,p}(\calT) 
        \diam(T)^{k} 
        \left( 
            \| \cartan u      \|_{L^{p}(T_{m})} 
            +
            \Cvier{n}(\calT)^{k+1} 
            \Cdetfive{n}(\calT)^{\frac 1 p} 
            \| \cartan u \|_{L^{p}(U_{m-1})}
        \right)
        \\&\qquad
        + 
        \Cvier{n}(\calT)^{k} 
        \Cdetfive{n}(\calT)^{\frac 1 p} 
        \| w \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
\end{theorem}

@@

    Let $u \in W^{p}\Alt^{k}(\Omega)$. 
    First, there exists $w_0 \in W^{p}\Alt^{k}(T_{0})$ satisfying $\cartan w_0 = \cartan u$ over $T_{0}$ together with 
    \begin{gather*}
        \| w_0 \|_{L^{p}(T_{0})} \leq C_{{\rm{PF}},k,T_{0},p} \| \cartan u \|_{L^{p}(T_{0})}.
    \end{gather*}
    %Suppose that claim is true when the domain has $m$ simplices of dimension $n$. 
    Suppose that $0 < m \leq M$ such that there exists $w_{m-1} \in W^{p}\Alt^{k}(\Omega_{m-1})$ 
    with $\cartan w_{m-1} = \cartan u$ over $\Omega_{m-1}$. 
    By assumption, $T_{m}$ and $\overline\Omega_{m-1}$ share the interface $\Gamma_{m}$, which is a collection of faces of $T_{m}$. 
    In accordance to Lemma~\ref{lemma:existenceofstar}, adding $T_{m}$ completes a star in $\calT$ around some simplex, and we let $U_{m-1} \subseteq \Omega_{m-1}$ be the complement of $T_{m}$ in that newly completed star.

@@

    We want to estimate norms. 
    By construction,
    \begin{align*}
        \| w_{m} \|_{L^{p}(T_{m})}
        \leq  
        \| w'_{m} \|_{L^{p}(T_{m})}
        + 
        \| w''_{m} \|_{L^{p}(T_{m})}
        .
    \end{align*}
    Due to Lemma~\ref{lemma:mixedbconsimplex:exteriorderivative},
    we can assume 
    \begin{align*}
        \| w''_{m} \|_{L^{p}(T_{m})} 
        &
        \leq 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} \| \cartan u - \cartan w'_{m} \|_{L^{p}(T_{m})}
        \\&
        \leq 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} \| \cartan u      \|_{L^{p}(T_{m})} 
        + 
        C_{{\rm PF},T_{m},\Gamma_{m},k,p} \| \cartan w'_{m} \|_{L^{p}(T_{m})} 
        .
    \end{align*}
    We also employ Theorem~\ref{theorem:pullbackestimate} to find 
    \begin{align*}
        \| w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        |\det(\Jacobian \Xi_{m}  )|^{-\frac 1 p} 
        \| \Jacobian \Xi_{m} \|^{k}
        \cdot 
        \| w_{m-1} \|_{L^{p}(U_{m-1})}
        ,
        \\
        \| \cartan w'_{m} \|_{L^{p}(T_{m})}
        &\leq 
        |\det(\Jacobian \Xi_{m}  )|^{-\frac 1 p} 
        \| \Jacobian \Xi_{m} \|^{k+1}
        \cdot 
        \| \cartan w_{m-1} \|_{L^{p}(U_{m-1})}
        .
    \end{align*}
    We have assumed that $\cartan w_{m-1} = \cartan u_{|\Omega_{m-1}}$. 
    We use the properties of the reflection mapping in Theorem~\ref{theorem:starreflection}. 
%     \begin{align*}
%         ...
%     \end{align*}
%     This enables the recursion
%     \begin{align*}
%         \| w_{m} \|_{L^{p}(T_{m})}
%         &
%         \leq  
%         |\det(\Jacobian \Xi_{m}  )|^{-\frac 1 p} 
%         \| \Jacobian \Xi_{m} \|^{k} 
%         \| w_{m-1} \|_{L^{p}(U_{m-1})}
%         \\&\qquad
%         + 
%         C_{{\rm PF},T_{m},\Gamma_{m},k,p} 
%         \| \cartan u      \|_{L^{p}(T_{m})} 
%         \\&\qquad
%         + 
%         C_{{\rm PF},T_{m},\Gamma_{m},k,p} 
%         |\det(\Jacobian \Xi_{m}  )|^{-\frac 1 p} 
%         \| \Jacobian \Xi_{m} \|^{k+1} 
%         \| \cartan u \|_{L^{p}(U_{m-1})}
%         .
%     \end{align*}
    The existence of $w \in W^{p}\Alt^{k}(\Omega)$ satisfying the recursive estimate follows. 
\end{proof}
